Follow-up Samples

### 2016-12-05 submission/manuscript_outline_20161024.Rmd
abstract changes

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this current study we attempt to combine specific physical characteristics, patient history, Fecal Immunochemical Test (FIT), and the bacterial microbiome to build the best model for classifying those with and without lesion (adenoma or carcinoma).  From this improved model we applied it to a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. Carcinoma dependent differences in initial and follow up samples were identified for the overall bacterial community composition (P-value = ```r paste(beta_summary['Carcinoma', 'PERMANOVA'])```).  The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  There was a significant decrease in probabliity of lesion from the model after treatment for both the adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```) group.  The top 5 most important operational taxonomic units for this model were ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  For prediction on samples with follow up the initial sensitivity was ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) and follow up sensitivity was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, this model can accurately predict and subsequently respond to treatment by lowering the probability of lesion.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this current study we attempt to combine specific physical characteristics, patient history, Fecal Immunochemical Test (FIT), and the bacterial microbiome to build the best model for classifying those with and without lesion (adenoma or carcinoma).  This model was tested on a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. Using NMDS, the thetayc distance between initial and follow up showed a significant difference (P-value = ```r paste(beta_summary['Carcinoma', 'PERMANOVA'])```) for the carcinoma group only.  The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  There was a significant decrease in probabliity of lesion from the model after treatment for both the adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```) group.  The top 5 most important operational taxonomic units for this model were ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  For prediction on samples with follow up the initial sensitivity was ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) and follow up sensitivity was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, this model can accurately predict and subsequently respond to treatment by lowering the probability of lesion.
